---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


















## Are there differences in activity patterns between weekdays and weekends?
### (Assuming that the data file activity.csv is in the working directory)

## Loading and preprocessing the data
```{r,echo=TRUE}
activity <- read.csv("activity.csv") 
```

## What is mean total number of steps taken per day?  

A histogram of the total number of steps taken each day is followed.  

```{r,echo=TRUE, results="asis", fig.height=4}
StepsPerDay <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)  
StepsPerDay <- as.data.frame(StepsPerDay)  
library(ggplot2)  
ggplot(StepsPerDay, aes(x=StepsPerDay)) + geom_histogram(binwidth=1000, colour="black", fill="white")  
```

With the following code, we could get the mean of the total number of steps taken per day.
```{r,echo=TRUE}
meanperday <- round(mean(StepsPerDay$StepsPerDay),2)  
meanperday  
medianperday <- median(StepsPerDay$StepsPerDay)  
medianperday  
```

The mean of the total number of steps taken per day is `r meanperday`. The median of the total number of steps taken per day is `r medianperday`.  

## What is the average daily activity pattern?

The time series plot of average daily activity pattern is followed.  

```{r,echo=TRUE} 
IntervalAverage <- tapply(activity$steps, factor(activity$interval), mean, na.rm = TRUE)  
IntervalAverage <- data.frame(Average=IntervalAverage, interval = as.numeric(names(IntervalAverage)))  
ggplot(IntervalAverage, aes(x=interval, y=Average)) + geom_line()  
```

We can observe the maximum numbers of steps in the plot, but when the maxumum occured is not easy to tell. We can get it using the following code.

```{r,echo=TRUE}
max_interval <- IntervalAverage[which(IntervalAverage$Average==max(IntervalAverage$Average)),"interval"];
max_interval  
max_hour <- max_interval%/%60  
max_min <- max_interval%%60  
max_time <- paste(max_hour,":", max_min, sep="")  
max_time  
```

On average across all the days in the dataset, the 5-minute interval started at `r max_interval`, i.e., `r max_time`, contains the maximum number of steps.  


## Imputing missing values

To check how many values are missing in the data set, we could sum up the return of is.na() function.  
```{r,echo=TRUE}
total_na <- sum(is.na(activity$steps))  
total_na  
```

There are `r total_na` missing values in the dataset.  


For these missing values, we used the mean for a 5-minute interval to fill a missing value in that interval.  

A histogram of the total number of steps taken each day in the imputed data set is followed.   

```{r,echo=TRUE}
activity_new <- activity  
activity_new$steps[is.na(activity_new$steps)] <- IntervalAverage$Average[match(activity_new$interval[is.na(activity_new$steps)], IntervalAverage$interval)]  
StepsPerDay_imputed <- tapply(activity_new$steps, activity_new$date, sum)  
StepsPerDay_imputed <- as.data.frame(StepsPerDay_imputed)  
ggplot(StepsPerDay_imputed, aes(x=StepsPerDay_imputed)) + geom_histogram(binwidth=1000, colour="black", fill="white")  
```
 
In the imputed data set, the mean of the total number of steps taken per day should be different from the mean in the original data set. 
```{r,echo=TRUE}
meanperday_imputed <- round(mean(StepsPerDay_imputed$StepsPerDay_imputed),2)  
meanperday_imputed  
medianperday_imputed <- round(median(StepsPerDay_imputed$StepsPerDay_imputed),2)  
medianperday_imputed  
```
In the imputed data set, the mean of the total number of steps taken per day is `r meanperday_imputed`; the median of the total number of steps taken per day is `r medianperday_imputed`. 

They are different from these estimated in the first part of the assignment. The mean and median of the total number of steps taken per day in the imputed data set are larger than these in the original data set. It is because in the previous data set, NAs contribute zero step to the total steps in that day, while imputed values contribute steps larger than zero. Therefore, mean increases and median tends to increase.


## Are there differences in activity patterns between weekdays and weekends?

The following panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r,echo=TRUE}
activity_new$day <- c("weekday", "weekend")[(weekdays(as.Date(activity_new$date)) %in% c("Saturday","Sunday")) + 1]  

activity_new_weekday <- activity_new[activity_new$day=="weekday",]  
activity_new_weekend <- activity_new[activity_new$day=="weekend",]  

activity_new_weekday_interval_average <- tapply(activity_new_weekday$steps, activity_new_weekday$interval,mean)  
interval1=names(activity_new_weekday_interval_average)  
activity_new_weekday_interval_average <- data.frame(Average=activity_new_weekday_interval_average,  day="weekday")  

activity_new_weekend_interval_average <- tapply(activity_new_weekend$steps, activity_new_weekend$interval,mean)  
interval2=names(activity_new_weekend_interval_average)  
activity_new_weekend_interval_average <- data.frame(Average=activity_new_weekend_interval_average, day="weekend")  

activity_new_weekdays_interval_average <- rbind(activity_new_weekday_interval_average, activity_new_weekend_interval_average)  
activity_new_weekdays_interval_average$interval <- as.numeric(c(interval1, interval2))  
ggplot(data=activity_new_weekdays_interval_average, aes(x=interval, y=Average))  + geom_line() + facet_wrap( ~ day, ncol=1)  
```
